# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r loading,cache=TRUE}
unzip('activity.zip')
raw <- read.csv('activity.csv',
                colClasses=c('numeric', 'character', 'numeric'),
                na.strings='NA')

# remove rows with NA values                
dat <- raw[!is.na(raw$steps),]
dat <- transform(dat, date = as.Date(date, format='%Y-%m-%d'))

```

## Mean total number of steps taken per day

```{r}
# total number of steps each day
stepsDate <- aggregate(steps~date, data=dat, sum)
hist(stepsDate$steps, main='Historgram of total steps each day')

# compute mean and median
stepsMean <- mean(stepsDate$steps)
stepsMedian <- median(stepsDate$steps)
```

Mean of total steps per day: **`r round(stepsMean, 2)`**.

Median of total steps per day: **`r round(stepsMedian, 2)`**.

## The average daily activity pattern

```{r,fig.width=11,fig.height=5}
stepsInterval <- aggregate(steps ~ interval, dat, mean)
plot(stepsInterval$interval, stepsInterval$steps, type='l',
     xlab='Interval', ylab='Average number of steps')

# find interval with maximum steps
maxRow <- which.max(stepsInterval[,'steps'])
maxInt <- stepsInterval[maxRow,'interval']
abline(v=maxInt,col=4,lty=3)
text(maxInt,0,maxInt,pos=4)
```

Maximum average **`r round(max(stepsInterval$steps), 2)`** steps in
interval **`r maxInt`**.

## Imputing missing values

```{r}
missing <- sum(is.na(raw$steps))
```

There are **`r missing`** intervals with missing values for number of steps.
Missing values will be replaced with an average number of available values.

```{r}
# replace missing values with an average value of available numbers  
avgSteps <- mean(dat$steps)
raw[is.na(raw)] <- avgSteps

fixedData <- aggregate(steps ~ date, data = raw, sum)
hist(fixedData$steps)

meanSteps <- mean(fixedData$steps)

medianSteps <- median(fixedData$steps)

```

Average number of steps after missing values fixed: **`r meanSteps`**.

Median number of steps after missing values fixed: **`r medianSteps`**.

## Differences in activity patterns between weekdays and weekends

```{r}
dat <- transform(raw, weekday=weekdays(as.Date(date, '%Y-%m-%d')))
dat$day <- as.factor(ifelse(dat$weekday == 'Saturday'
                                | dat$weekday == 'Sunday', 'weekend', 'weekday'))

stepsWd <- aggregate(steps ~ interval + day, dat, mean)

library(lattice)
xyplot(steps ~ interval | day, stepsWd, type='l', layout=c(1,2))
```
